Are populations like a circuit? The relationship between isolation by distance and isolation by resistance

### 2018-10-23 isolation_by_coalescence.tex
final!! pass

\begin{abstract}
A number of methods commonly used in landscape genetics
use an analogy to electrical resistance on a network
to describe and fit barriers to movement across the landscape
using genetic distance data.
These are motivated by a mathematical equivalence between electrical resistance 
between two nodes of a network
and the ``commute time'',
which is the mean time for a random walk on that network to leave one node, visit the other, and return.
However, genetic data are more accurately modeled by a different quantity,
the coalescence time.
Here, we describe the differences between resistance distance and coalescence time,
and explore the consequences for inference.
We implement a Bayesian method to infer effective movement rates and population sizes 
under both these models,
and find that inference using commute times
can dramatically mislead in the presence of biased gene flow.
We then use forwards-time simulation with continuous geography to demonstrate that
coalescence-based inference remains more accurate than resistance-based methods on realistic data,
but difficulties highlight the need for methods 
that explicitly model continuous, heterogeneous geography.
\end{abstract}

@@

As motivated by \citet{mcrae2006isolation},
commute time (i.e., resistance distance) is a computationally tractable approximation
to the more accurate random walk-based model of genetic distance,
the \emph{coalescence time}.
Genetic distances between two genomes sampled from the landscape
do derive from an average across a large number of paths between those points --
the lineages along which each segment of genetic material has been inherited
from their most recent common ancestor.
The more recent this ancestor is,
the smaller the genetic distance will be.
Thanks to recombination, there are a large number of such paths, 
and genetic distance averages over these.
It is reasonable in some situations to model these lineages as random walks across the landscape;
however, the model of genetic distance that one is thus led to
is \emph{coalescence time} of the random walks, 
rather than commute time, the quantity that corresponds to resistance.
This naturally raises the question: 
Are methods that depend on effective resistances being misled by model misspecification?
If so, would using coalescence time do better?

\citet{mcrae2006isolation} showed that coalescence time and resistance distance are equal
(up to shift and scaling) in isotropic landscapes such as a ring \citep{matsen2006convergence},
and found that the two were highly correlated in several test cases.
However, all test cases had symmetric migration
(the same rate of gene flow in both directions along each edge),
which is likely not the case in many situations, e.g.,
in rivers \citep{morrissey2009maintenance,sundqvist2016directional,hanks2017modeling},
source-sink dynamics \citep{dias1996sources},
or after population expansions \citep{with2002landscape}.

In this paper, we contrast 
coalescence time and resistance distance,
and develop an inference method for movement rates on a discrete landscape
using both as estimated from genetic distances.
As we will see,
although the two methods are conceptually similar,
resistance-based inference may get the wrong answer if data are derived from a coalescent process,
especially in the presence of gene flow asymmetry.
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{shaffer2017desert}.
Other methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information;
however, the geographical resolution is much coarser.
(A recently-released method 
that is based on EEMS but uses haplotype length data and an underlying model of coalescence time
\citep{alasadi2018estimating}
is promising, but still produces similarly coarse maps.)
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem,
both because of unmodeled ``process noise'' involved in discretization of geographical populations,
and the mathematical structure of the underlying problem.

@@

\begin{figure}
\centering
     \includegraphics[scale=.5]{figs/conceptn}
    \caption{
    Illustration of the conceptual differences
    betweeen coalescence and commute time between states 1 and 3 
    for a continuous time Markov chain with four linearly connected states. 
    Coalescence time (left) is the time for two independently moving particles to meet and coalesce 
    (it is possible for them to be in the same location without coalescing).
    Commute time is time for a single particle starting in one state to travel to another state 
    and then back to the original state.
    Note how the time axis is scaled to be faster in the figure showing commute time
    than the one for coalescence time in order for them to fit in the same space.
    } \label{fig:concept_coalcom}
\end{figure}

@@

Now, we explain how to compute the relevant quantities --
mean commute and coalescence times -- from a Markov chain model of lineage movement.
The Markov chain is defined by its \emph{generator matrix}, denoted $G$,
for which $G_{xy}$ gives the jump rate from $x$ to $y$.
This means that if the Markov chain is at location $X_t$ at time $t$,
then for $x \neq y$, the probability the chain jumps from $x$ to $y$ in the next unit of time
is proportional to $G_{xy}$,
i.e., $\PP{X_{t+\epsilon} = y \given X_t = y} = \epsilon G_{xy} + O(\epsilon^2)$,
and $G_{xx} = - \sum_{y \neq x} G_{xy}$.
We will need to find the \emph{hitting times} of the chain --
i.e., for each pair of states $x$ and $y$, 
the mean time until the chain first hits $y$ after being started at $x$.
We denote this quantity $H_{xy}$.
Throughout, we assume that all hitting times are finite,
which implies the chain is connected and irreducible.
By conditioning on $X_\epsilon$,
one can use the Markov property to show that
\begin{align} \label{eqn:hitting_sum}
    \sum_y G_{xy} H_{yz} = -1 \qquad \text{for} \; x \neq z,
\end{align}
i.e., if $G_{-z}$ is the matrix with the $z^\text{th}$ row and column removed,
and $\bone$ is the vector composed of all ones,
then the $z^\text{th}$ column of $H$, except for $H_{zz}$,
can be computed as $- (G_{-z})^{-1} \bone$.
This, along with $H_{zz} = 0$, allows computation of $H$
from the movement rates $G$.

@@

This suggests that nonuniqueness of solutions may not be a problem in practice,
as long as geography is discretized into sufficiently many locations
and direct, long-distance migration is disallowed.
However, another problem appears at finer geographic resolutions:
although solutions may be in principle unique,
the problems become \emph{ill-conditioned} 
in the sense that arbitrarily small variations in the input times
(even numerical instability)
may produce large differences in the inferred rates.
Roughly speaking, this occurs because as the spatial resolution increases,
the matrix $G$ converges to a second-order elliptic differential operator \citep{stroock1997multidimensional};
such operators are deformations of the Laplacian \citep{feller4},
which is well-known to have eigenvalues that decay rapidly 
\citep{hpmckean1967curvature,kuttler1984eigenvalues}.
Information about the landscape contained in higher-order modes
(e.g., finer resolution changes in hitting times)
is then almost obliterated by the small eigenvalues of $G$,
making the inverse problem possible in theory but not in practice.
For more discussion of related problems, see 
\citet{epstein2008badtruth}, \citet{myers2008learn}, or \citet{terhorst2015fundamental}.

@@

Therefore, we use Bayesian inference,
placing independent exponential priors on the parameters --
the movement rates, $G_{ij}$, that are not constrained to be zero
and the coalescence rates, $\gamma$.
Concretely, suppose that the spatial arrangement of populations is given by a graph,
and only movement rates corresponding to edges in the graph are allowed to be nonzero.
Write $i \sim j$ to mean that $i$ and $j$ are adjacent in the graph,
that there are $m_G$ edges in the graph,
and that each of the $n$ populations has its own coalescence rate.
Then, the resulting log-posterior density is
\begin{align} \label{eq:post}
    \begin{split}
\mathcal{L}(D, G, \gamma) 
    &=
    2 m_G \log(\lambda_G) + n \log(\lambda_\gamma) 
    - \frac{n(n+1)}{2} \log(2 \pi \sigma^2_\epsilon)
	- \sum_{i \sim j} \lambda_G G_{ij}  \\
    &\qquad
    -\sum_{i=1}^n \lambda_{\gamma}\gamma_i
		-\frac{1}{2 \sigma_{\epsilon}^2} \sum_{i \leq j} \left(
            \mathcal{C}_{ij}(G,\gamma) - D_{ij}
        \right)^2 
    \end{split}
\end{align}
Here $1/\lambda_\gamma$ and $1/\lambda_G$ are the prior means of $\gamma_i$ and $G_{ij}$,
respectively.

@@

\paragraph{Forwards-time simulations}
To produce realistic data, we implemented forwards-time simulations using SLiM v3.1 
\citep{haller2018forward,haller2018treesequence},
with individuals living across continuous, two-dimensional geography (sometimes with barriers),
from which we recorded genomic data.
The basic simulation, which we modified to produce several other situations,
is as follows.
Each simulation had around 6400 diploid, hermaphroditic individuals, 
living across two-dimensional geography.
Mates were selected from nearby individuals using a Gaussian kernel with standard deviation $\sigma_d$,
truncated at $3 \sigma_d$.
Simulations were done on either a square region with width equal to $80\sigma_d$,
or a $5:3$ rectangular region with the same area.
Offspring are dispersed away from the mother using the same kernel,
reflected from the boundaries of the habitable area.
To produce a realistic model and prevent excessive population clumpiness \citep{felsensten1975pain}, 
we model local competition instead of a globally constant population size.
To do this, 
we use the same truncated Gaussian kernel to define an ``interaction strength''
between each pair of nearby individuals, 
and compute the probability of survival of each individual to the next time step as
$\min(0.9, 2 / (1 +  5x/12 \pi))$,
where $x$ is the sum of the interaction strengths of that individual with all other individuals
within distance $3 \sigma_d$.
(The value $x$ is an estimate of local population density.)
This reduces fitness for individuals in denser areas,
and produces an equilibrium population density
of close to one individual per $\sigma_d^2$ units of area.
The probability of survival for individuals within $10\sigma_d$ of the edge of the habitat
was multiplied by a linear factor that scaled from 0 to 1 with distance from the edge.
Configuration the resulting simulations was similar to what is seen in 
Figure \ref{fig:5x3b_continuous}.

@@

We first tested the method using data generated under the model.
To test the methods across a wide range of situations,
we generated fifty sets of random movement rates on $3 \times 2$ rectangular grids:
twenty-five symmetric (i.e., with $G_{ij} = G_{ji}$) and twenty-five asymmetric.
We generated movement parameters
by rounding independent exponential random variables up to the nearest one tenth,
and fixed coalescence rates at 1 for all locations.
We then analytically computed mean coalescence times,
and added independent noise to each pairwise coalescence time.
We assume that we can estimate the amount of noise reasonably well, 
so used the true standard deviation of the noise for $\sigma_\epsilon$ in the likelihood function.
To assess the impact on inference of the number of parameters,
we performed inference both with and without the assumption 
that coalescence rate is the same everywhere.
For $3 \times 2$ graphs, there are 21 equations
and 14 movement parameters,
so that with a global coalescence rate there are 15 unknowns
but with individual coalescence rates there are 20 unknowns,
which is close to nonidentifiable.
To quantify accuracy across these random graphs in different situations,
we computed, for each model fit on a particular random graph,
the absolute difference between the posterior median of each movement parameter
and the true value.
We then averaged these across movement parameters to produce a measure of accuracy for that model fit,
reported below as ``mean absolute error''.

\paragraph{Varying noise}
Unsurprisingly, coalescence time inference became less accurate as the amount of noise increased.
As we varied the standard deviation of the noise
from 1/1000 to 1/50 of the mean value of $C$ for that grid,
the mean absolute error of inferred gene flow rates ($G_{ij}$) 
for coalescence time inference with a single coalescence rate
increased from around 0.05 to 0.4.
Since true values of $G_{ij}$ were of order 1,
this is a transition from precise to very rough (but still informative) inference.
Posterior interquartile ranges increased proportionally,
although there was substantial variation in accuracy between replicates;
see Figures \ref{fig:mult_noise} and \ref{fig:mult_noise_iqr}.
Increasing the number of parameters by allowing multiple coalescence rates
made the inference problem even harder --
mean absolute errors no longer depended strongly on the amount of noise added,
and hovered around 0.4.

@@

\paragraph{Varying coalescence rates}
Lower coalescence rates mean that lineages wander about the landscape for longer
before they coalesce, 
thus greatly reducing the geographic information we get from coalescence times.
Indeed, coalescence rates had a strong effect on feasibility of the inference problem
(shown in Figures \ref{fig:mult_gam} and \ref{fig:mult_gam_iqr}):
lowering the coalescence rate from 10 to 0.1
(always adding noise with standard deviation 1/200th of the mean value)
increased mean absolute error of coalescence time inference from around 0.05 to 0.5.
Again, allowing multiple coalescence rates
or using resistance distance inference were highly inaccurate in all cases,
with mean absolute errors of above 0.5.

@@

We then test the method's ability to locate barriers to migration
on the $5 \times 3$ grid with asymmetric gene flow shown in Figure \ref{fig:5x3b_grid}. 
Each nonzero movement rate is determined randomly as before,
and coalescence rates are set to 1 everywhere.
Figure \ref{fig:5x3b_post_coalvcom} shows posterior distributions of the movement rates
for both coalescence time and resistance distance inference
on this graph, with noise standard deviation equal to
1/1000 of the mean coalescence time (left)
and 1/100 of the mean coalescence time (right).
Although increased noise in the data increased uncertainty in the estimated values,
coalescence time inference correctly inferred not only the location of the barrier
but also the migration rates along all other edges within a reasonable margin of error.
Resistance distance inference also identified the barrier,
but was much less accurate for the other values.
We also explored making the problem harder,
by first (a) dropping data corresponding to populations 2, 5, 7, and 15;
and then (b) allowing a separate coalescence rate for each location,
in both cases using a noise level of 1/100.
The resulting posterior distributions are shown in Figure \ref{fig:5x3boxplots_mult_gam},
and show substantially increased error,
but only in movement rates to the removed locations.
As before, uncertainty was increased when multiple coalescence rates were allowed,
but not as strongly as before, likely because the problem is less close to being underconstrained,
having 105 equations and 59 unknowns.

@@

Both matrices of coalescence and commute times, $C$ and $R$, are symmetric,
but they do not deal with asymmetric (biased) migration in the same way.
Since commute time to location $i$ requires paths both to and from $i$,
but coalescence time to $i$ only requires lineages to leave $i$,
if there is a low rate of migration back into $i$,
then commute times to $i$ could be quite long even while coalescence times are not.
For instance, suppose there are three populations arranged in a line,
and lineages currently in the outer populations are much more likely to come from the central population
than the other way around.
Since lineages quickly move to the center and coalesce regardless of starting position,
coalescence time (and genetic diversity) will be relatively low between all individuals.
Commute time between the two outer locations, on the other hand,
will be much longer than other comparisons, since it requires a lineage to leave the center.

@@

Differences between true mean coalescence times ($C$) and mean resistance distances ($\comdist$) 
are shown in Figure \ref{fig:RvsC}:
they are fairly small in the uniform and symmetric cases, moderate in the asymmetric case, 
and extreme in the biased asymmetric case.
This difference suggests that resistance distance inference may be strongly misled 
in the asymmetric and biased situations,
but in principle it still could be the case that 
the movement rates that give the best fit of resistance distance
to coalescence time data could be close to the actual rates use to generate the data.
This does not seem to occur in practice:
coalescence time inference was substantially more accurate in all cases.
Posterior distributions of movement parameters are shown in Figure \ref{fig:4x4box},
produced by applying both inference methods with a noise level of 1/200.
The mean absolute errors for coalescence time inference are 0.12 or less in all cases.
Resistance distance inference obtains roughly uniform migration rates
(although varying by a factor of 2) in the uniform case;
and migration rates noisily correlated with the truth in the symmetric graph.
However, resistance distance inferences for the asymmetric and biased graphs are only weakly correlated, 
if at all, to the truth.
Resistance distance inference also drastically overestimates movement rates for the biased case,
as we would expect based on the differences discussed above.

@@

The data we have used thus far are produced ``under the model'',
and should provide an accurate depiction of our method applied to discrete, randomly mating populations
whose connections by migration are known.
However, this is nearly always a rough approximation to reality,
in which organisms are distributed across continuous geography,
and ``populations'' are constructed by necessity, often driven by sampling locations.
For this reason, we applied coalescence and resistance inference,
as well as the popular resistance-based method EEMS \citep{petkova2016visualizing},
to data from simulations on continuous geography.

\paragraph{Process noise}
The mean coalescence times of lineages that we compute analytically
are exact for large, randomly mating populations,
but the word ``mean'' indicates an average across several levels of randomness:
each observed, empirical mean estimated using a sample of genotypes
will deviate from this depending on both the individuals sampled (``sampling noise'')
\citep{ashander2018demographic},
and the stochasticity of the population history itself (``process noise'')
\citep{wakeley2012genealogies,waples2009modelling}.
% Both sources of noise can be substantially larger 
% in populations living across continuous geography
% than in discrete populations
% due to randomness in spatial locations.
% This is hinted at by Figure \ref{fig:5x3b_continuous},
% where we see that the distribution of individuals across the landscape is fairly patchy:
% a barrier might be inferred in a region that has not happened to have a local population recently
% only because of demographic stochasticity.
% This might make inference using a discretization of continuous space considerably more difficult.

To quantify the relative contributions of the different sources of noise,
we compared mean genetic distances, calculated in the same way,
between two nonoverlapping sets of samples from each of three identical simulations.
These simulations were done using SLiM as described in the Methods.
We found that with these parameter values, process and sampling noise were of similar magnitude,
contributing 6.7\% and 13.8\% of the variance, respectively
(Supplemental Figure \ref{fig:h_val_comp}).
% An implication of this is that uncertainties in genetic distance
% estimated from the data will be (possibly large) underestimates,
% a fact also pointed out for effective population size by \citet{waples2009modelling}.

@@

Our new method infers effective population sizes and gene flow rates quite well 
-- and, being Bayesian, provides estimates of uncertainty --
given data from discrete populations connected by migration.
Can our method replace resistance-based methods?
Perhaps, but
the substantial uncertainty we saw on graphs with only tens of nodes
is indicative of a larger problem we face for realistic models.
We have seen that discretization of space results in substantial modeling error,
in part because of randomness of geographic sampling
and unmodeled process noise can lead to overfitting.
Partitioning space into a finer grid should help with these problems,
but tends to make the inference problem itself more ill-conditioned:
with more connections,
changing the value of one connection affects coalescence time less,
and so inferences about that value must necessarily be less certain.
This tradeoff implies a high degree of unavoidable uncertainty.
Reproducible, reliable inference will likely require development of new inference methods
explicitly incorporating continuous geography.

>>>>

\begin{abstract}
A number of methods commonly used in landscape genetics
use an analogy to electrical resistance on a network
to describe and fit barriers to movement across the landscape
using genetic distance data.
These are motivated by a mathematical equivalence between electrical resistance 
between two nodes of a network
and the ``commute time'',
which is the mean time for a random walk on that network to leave one node, visit the other, and return.
However, genetic data are more accurately modeled by a different quantity,
the coalescence time.
Here, we describe the differences between resistance distance and coalescence time,
and explore the consequences for inference.
We implement a Bayesian method to infer effective movement rates and population sizes 
under both these models,
and find that inference using commute times
can produce misleading results in the presence of biased gene flow.
We then use forwards-time simulation with continuous geography to demonstrate that
coalescence-based inference remains more accurate than resistance-based methods on realistic data,
but difficulties highlight the need for methods 
that explicitly model continuous, heterogeneous geography.
\end{abstract}

@@

As motivated by \citet{mcrae2006isolation},
commute time (i.e., resistance distance) is a computationally tractable approximation
to the more accurate random walk-based model of genetic distance,
the \emph{coalescence time}.
Genetic distances between two genomes sampled from the landscape
do derive from an average across a large number of paths between those points --
the lineages along which each segment of genetic material has been inherited
from their most recent common ancestor.
Thanks to recombination, there are a large number of such paths, 
and genetic distance averages over these.
The more recent these common ancestors tend to be,
the smaller the genetic distance is.
It is reasonable in some situations to model these lineages as random walks across the landscape;
however, the model of genetic distance that one is thus led to
is \emph{coalescence time} of the random walks, 
rather than commute time, the quantity that corresponds to resistance.
This naturally raises the question: 
Are methods that depend on effective resistances being misled by model misspecification?
If so, would using coalescence time do better?

\citet{mcrae2006isolation} showed that coalescence time and resistance distance are equal
(up to shift and scaling) in isotropic landscapes such as a ring \citep{matsen2006convergence},
and found that the two were highly correlated in several test cases.
However, all test cases had symmetric migration
(the same rate of gene flow in both directions along each edge),
which is likely not the case in many situations, e.g.,
in rivers \citep{morrissey2009maintenance,sundqvist2016directional,hanks2017modeling},
with source-sink dynamics \citep{dias1996sources},
or after population expansions \citep{with2002landscape}.

In this paper, we contrast 
coalescence time and resistance distance,
and develop a method to infer movement rates on a discrete landscape
using genetic distances.
As we will see,
although the two methods are conceptually similar,
resistance-based inference may get the wrong answer if data are derived from a coalescent process,
especially in the presence of gene flow asymmetry.
We also explore the question of identifiability.
Resistance-based methods are often used to predict distances based on a few layers of geographical data
(land cover, slope, etcetera).
Some methods find the combination of landscape layers to best explain genetic distance,
which allows very fine geographic resolution \citep{shaffer2017desert}.
Other methods such as EEMS \citep{petkova2016visualizing} infer conductances without such prior information;
however, the geographical resolution is much coarser.
(A recently-released method 
that is based on EEMS but uses haplotype length data and an underlying model of coalescence time
\citep{alasadi2018estimating}
is promising, but still produces similarly coarse maps.)
One reason for this difference in resolution
is that EEMS uses more computationally intensive Bayesian methods.
However, we show that this is a more general problem,
both because of unmodeled ``process noise'' involved in discretization of geographical populations,
and the mathematical structure of the underlying problem.

@@

\begin{figure}
\centering
     \includegraphics[scale=.5]{figs/conceptn}
    \caption{
    Illustration of the conceptual differences
    between coalescence and commute time between states 1 and 3 
    for a continuous time Markov chain with four linearly connected states. 
    Coalescence time (left) is the time for two independently moving particles to meet and coalesce 
    (it is possible for them to be in the same location without coalescing).
    Commute time is time for a single particle starting in one state to travel to another state 
    and then back to the original state.
    Note how the time axis is scaled to be faster in the figure showing commute time
    than the one for coalescence time in order for them to fit in the same space.
    } \label{fig:concept_coalcom}
\end{figure}

@@

Now, we explain how to compute the relevant quantities --
mean commute and coalescence times -- from a Markov chain model of lineage movement.
The Markov chain is defined by its \emph{generator matrix}, denoted $G$,
for which $G_{xy}$ gives the jump rate from $x$ to $y$.
This means that if the Markov chain is at location $X_t$ at time $t$,
then for $x \neq y$, the probability the chain jumps from $x$ to $y$ in the next unit of time
is proportional to $G_{xy}$,
i.e., $\PP{X_{t+\epsilon} = y \given X_t = y} = \epsilon G_{xy} + O(\epsilon^2)$,
and $G_{xx} = - \sum_{y \neq x} G_{xy}$.
We will need to find the \emph{hitting times} of the chain --
i.e., for each pair of states $x$ and $y$, 
the mean time until the chain first hits $y$ after being started at $x$.
We denote this quantity $H_{xy}$.
Throughout, we assume that all hitting times are finite,
which implies the chain is connected and irreducible.
By conditioning on $X_\epsilon$,
one can use the Markov property to show that
\begin{align} \label{eqn:hitting_sum}
    \sum_y G_{xy} H_{yz} = -1 \qquad \text{for} \; x \neq z,
\end{align}
i.e., if $G_{-z}$ is the matrix with the $z^\text{th}$ row and column removed,
and $\bone$ is the vector composed of all ones,
then the $z^\text{th}$ column of $H$, except for $H_{zz}$,
can be computed as $- (G_{-z})^{-1} \bone$.
This, along with $H_{zz} = 0$, allows computation of $H$
from the movement rates $G$ \citep{kemeny1983finite}.

@@

This suggests that nonuniqueness of solutions may not be a problem in practice,
as long as geography is discretized into sufficiently many locations
and direct, long-distance migration is disallowed.
However, another problem appears at finer geographic resolutions:
although solutions may be in principle unique,
the problems become \emph{ill-conditioned} 
in the sense that arbitrarily small variations in the input times
(even numerical instability)
may produce large differences in the inferred rates.
Roughly speaking, this occurs because as the spatial resolution increases,
the matrix $G$ converges to a second-order elliptic differential operator \citep{stroock1997multidimensional};
such operators are deformations of the Laplacian \citep{feller4},
which is well-known to have eigenvalues that decay rapidly 
\citep{hpmckean1967curvature,kuttler1984eigenvalues}.
Information about the landscape contained in higher-order modes
(e.g., finer resolution changes in hitting times)
is then almost obliterated by the small eigenvalues of $G$,
making the inverse problem possible in theory but not in practice.
This implies a fundamental limit to the geographical resolution of inferred maps.
For more discussion of related problems, see 
\citet{epstein2008badtruth}, \citet{myers2008learn}, or \citet{terhorst2015fundamental}.

@@

Therefore, we use Bayesian inference,
placing independent exponential priors on the parameters --
the movement rates, $G_{ij}$, that are not constrained to be zero
and the coalescence rates, $\gamma$.
Concretely, suppose that the spatial arrangement of populations is given by a graph,
and only movement rates corresponding to edges in the graph are allowed to be nonzero.
Suppose there are $m_G$ edges in the graph,
that each of the $n$ populations has its own coalescence rate,
and write $i \sim j$ to mean that $i$ and $j$ are adjacent in the graph.
Then, the resulting log-posterior density is
\begin{align} \label{eq:post}
    \begin{split}
\mathcal{L}(D, G, \gamma) 
    &=
    2 m_G \log(\lambda_G) + n \log(\lambda_\gamma) 
    - \frac{n(n+1)}{2} \log(2 \pi \sigma^2_\epsilon)
	- \sum_{i \sim j} \lambda_G G_{ij}  \\
    &\qquad
    -\sum_{i=1}^n \lambda_{\gamma}\gamma_i
		-\frac{1}{2 \sigma_{\epsilon}^2} \sum_{i \leq j} \left(
            \mathcal{C}_{ij}(G,\gamma) - D_{ij}
        \right)^2 
    \end{split}
\end{align}
Here $1/\lambda_\gamma$ and $1/\lambda_G$ are the prior means of $\gamma_i$ and $G_{ij}$,
respectively.

@@

\paragraph{Forwards-time simulations}
To produce realistic data, we implemented forwards-time simulations using SLiM v3.1 
\citep{haller2018forward,haller2018treesequence},
with individuals living across continuous, two-dimensional geography (sometimes with barriers),
from which we recorded genomic data.
The basic simulation, which we modified to produce several other situations,
is as follows.
Each simulation had around 6400 diploid, hermaphroditic individuals, 
living across two-dimensional geography.
Mates were selected from nearby individuals using a Gaussian kernel with standard deviation $\sigma_d$,
truncated at $3 \sigma_d$.
Simulations were done on either a square region with width equal to $80\sigma_d$,
or a $5:3$ rectangular region with the same area.
Offspring are dispersed away from the mother using the same kernel,
reflected from the boundaries of the habitable area.
To produce a realistic model and prevent excessive population clumpiness \citep{felsensten1975pain}, 
we model local competition instead of a globally constant population size.
To do this, 
we use the same truncated Gaussian kernel to define an ``interaction strength''
between each pair of nearby individuals, 
and compute the probability of survival of each individual to the next time step as
$\min(0.9, 2 / (1 +  5x/12 \pi))$,
where $x$ is the sum of the interaction strengths of that individual with all other individuals
within distance $3 \sigma_d$.
(The value $x$ is an estimate of local population density.)
This reduces fitness for individuals in denser areas,
and produces an equilibrium population density
of close to one individual per $\sigma_d^2$ units of area.
The probability of survival for individuals within $10\sigma_d$ of the edge of the habitat
was multiplied by a linear factor that scaled from 0 to 1 with distance from the edge.
Patchiness of the resulting simulations was similar to what is seen in 
Figure \ref{fig:5x3b_continuous}.

@@

We first tested the method using data generated under the model.
To test the methods across a wide range of situations,
we generated fifty sets of random movement rates on $3 \times 2$ rectangular grids:
twenty-five symmetric (i.e., with $G_{ij} = G_{ji}$) and twenty-five asymmetric.
We generated movement parameters
by rounding independent exponential random variables up to the nearest one tenth,
and fixed coalescence rates at 1 for all locations.
We then analytically computed mean coalescence times,
and added independent noise to each pairwise coalescence time.
We assume that we can estimate the amount of noise reasonably well, 
so used the true standard deviation of the noise for $\sigma_\epsilon$ in the likelihood function.
To assess the impact on inference of the number of parameters,
we performed inference both with and without the assumption 
that coalescence rate is the same everywhere.
For $3 \times 2$ graphs, there are 21 equations
and 14 movement parameters,
so that with a global coalescence rate there are 15 unknowns.
However, with individual coalescence rates there are 20 unknowns,
which is close to nonidentifiable.
To quantify accuracy across these random graphs in different situations,
we computed, for each model fit on a particular random graph,
the absolute difference between the posterior median of each movement parameter
and the true value.
We then averaged these across movement parameters to produce a measure of accuracy for that model fit,
reported below as ``mean absolute error''.

\paragraph{Varying noise}
Unsurprisingly, coalescence time inference became less accurate as the amount of noise increased.
As we varied the standard deviation of the noise
from 1/1000${}^\text{th}$ to 1/50${}^\text{th}$ of the mean value of $C$ for that grid,
the mean absolute error of inferred gene flow rates ($G_{ij}$) 
for coalescence time inference with a single coalescence rate
increased from around 0.05 to 0.4.
Since true values of $G_{ij}$ were of order 1,
this is a transition from precise to very rough (but still informative) inference.
Posterior interquartile ranges increased proportionally,
although there was substantial variation in accuracy between replicates;
see Figures \ref{fig:mult_noise} and \ref{fig:mult_noise_iqr}.
Increasing the number of parameters by allowing multiple coalescence rates
made the inference problem much harder --
mean absolute errors no longer depended strongly on the amount of noise added,
and hovered around 0.4.

@@

\paragraph{Varying coalescence rates}
Lower coalescence rates mean that lineages wander about the landscape for longer
before they coalesce, 
thus greatly reducing the geographic information we get from coalescence times 
\citep{wilkins2004separationoftimescales}.
Indeed, coalescence rates had a strong effect on feasibility of the inference problem
(shown in Figures \ref{fig:mult_gam} and \ref{fig:mult_gam_iqr}):
lowering the coalescence rate from 10 to 0.1
(always adding noise with standard deviation 1/200$^\text{th}$ of the mean value)
increased mean absolute error of coalescence time inference from around 0.05 to 0.5.
Again, allowing multiple coalescence rates
or using resistance distance inference were highly inaccurate in all cases,
with mean absolute errors of above 0.5.

@@

We then test the method's ability to locate barriers to migration
on the $5 \times 3$ grid with asymmetric gene flow shown in Figure \ref{fig:5x3b_grid}. 
Each nonzero movement rate is determined randomly as before,
and coalescence rates are set to 1 everywhere.
Figure \ref{fig:5x3b_post_coalvcom} shows posterior distributions of the movement rates
for both coalescence time and resistance distance inference
on this graph, with noise standard deviation equal to
1/1000${}^\text{th}$ of the mean coalescence time (left)
and 1/100${}^\text{th}$ of the mean coalescence time (right).
Although increased noise in the data increased uncertainty in the estimated values,
coalescence time inference correctly inferred not only the location of the barrier
but also the migration rates along all other edges within a reasonable margin of error.
Resistance distance inference also identified the barrier,
but was much less accurate for the other values.
We also explored making the problem harder,
by first (a) dropping data corresponding to populations 2, 5, 7, and 15;
and then (b) allowing a separate coalescence rate for each location,
in both cases using a noise level of 1/100.
The resulting posterior distributions are shown in Figure \ref{fig:5x3boxplots_mult_gam},
and show substantially increased error,
but only in movement rates to the removed locations.
As before, uncertainty was increased when multiple coalescence rates were allowed,
but not as strongly as before, likely because the problem is less close to being underconstrained,
having 105 equations and 59 unknowns.

@@

Both matrices of coalescence and commute times, $C$ and $R$, are symmetric,
but they do not deal with asymmetric (biased) migration in the same way.
Since commute time to location $i$ requires paths both to and from $i$,
but coalescence time to $i$ does not,
if there is a low rate of migration back into $i$,
then commute times to $i$ could be quite long even while coalescence times are not.
For instance, suppose there are three populations arranged in a line,
and lineages currently in the outer populations are much more likely to come from the central population
than the other way around.
Since lineages quickly move to the center and coalesce regardless of starting position,
coalescence time (and genetic diversity) will be relatively low between all individuals.
Commute time between the two outer locations, on the other hand,
will be much longer than other comparisons, since it requires a lineage to leave the center.

@@

Differences between true mean coalescence times ($C$) and mean resistance distances ($\comdist$) 
are shown in Figure \ref{fig:RvsC}:
they are fairly small in the uniform and symmetric cases, moderate in the asymmetric case, 
and extreme in the biased asymmetric case.
This difference suggests that resistance distance inference may be strongly misled 
in the asymmetric and biased situations,
but in principle it still could be the case that 
the movement rates that give the best fit of resistance distance
to coalescence time data could be close to the actual rates use to generate the data.
This does not seem to occur in practice:
coalescence time inference was substantially more accurate in all cases.
Posterior distributions of inferred movement parameters are shown in Figure \ref{fig:4x4box}.
The mean absolute errors for coalescence time inference are 0.12 or less in all cases.
Resistance distance inference obtains roughly uniform migration rates
(although varying by a factor of 2) in the uniform case;
and migration rates noisily correlated with the truth in the symmetric graph.
However, resistance distance inferences for the asymmetric and biased graphs are only weakly correlated, 
if at all, to the truth.
Resistance distance inference also drastically overestimates movement rates for the biased case,
as we would expect based on the differences discussed above.

@@

The data we have used thus far are produced ``under the model'',
and should provide an accurate depiction of our method applied to discrete, randomly mating populations
whose connections by migration are known.
However, this is nearly always a rough approximation to reality,
in which organisms are distributed across continuous geography,
and ``populations'' are constructed by necessity, often driven by sampling locations.
For this reason, we applied coalescence and resistance inference,
as well as the resistance-based method EEMS \citep{petkova2016visualizing},
to data from simulations on continuous geography.

\paragraph{Process noise}
The mean coalescence times of lineages that we compute analytically
are exact for large, randomly mating populations,
but the word ``mean'' indicates an average across several levels of randomness:
each observed, empirical mean estimated using a sample of genotypes
will deviate from this depending on both the individuals sampled (``sampling noise'')
\citep{ashander2018demographic},
and the stochasticity of the population history itself (``process noise'')
\citep{wakeley2012genealogies,waples2009modelling}.
% Both sources of noise can be substantially larger 
% in populations living across continuous geography
% than in discrete populations
% due to randomness in spatial locations.
% This is hinted at by Figure \ref{fig:5x3b_continuous},
% where we see that the distribution of individuals across the landscape is fairly patchy:
% a barrier might be inferred in a region that has not happened to have a local population recently
% only because of demographic stochasticity.
% This might make inference using a discretization of continuous space considerably more difficult.
To quantify the relative contributions of the different sources of noise,
we compared mean genetic distances, calculated in the same way,
between two nonoverlapping sets of samples from each of three identical simulations.
These simulations were done using SLiM as described in the Methods.
We found that with these parameter values, process and sampling noise were of similar magnitude,
contributing 6.7\% and 13.8\% of the variance, respectively
(Supplemental Figure \ref{fig:h_val_comp}).
% An implication of this is that uncertainties in genetic distance
% estimated from the data will be (possibly large) underestimates,
% a fact also pointed out for effective population size by \citet{waples2009modelling}.

@@

Our new method infers effective population sizes and gene flow rates quite well 
-- and, being Bayesian, provides estimates of uncertainty --
given data from discrete populations connected by migration.
Can our method replace resistance-based methods?
Perhaps, but
the substantial uncertainty we saw on graphs with only tens of nodes
is indicative of a larger problem we face for realistic models.
We have seen that discretization of space results in substantial modeling error,
in part because of randomness of geographic sampling
and unmodeled process noise can lead to overfitting.
Partitioning space into a finer grid should help with these problems,
but tends to make the inference problem itself more ill-conditioned:
with more connections,
changing the value of one connection affects coalescence time less,
and so inferences about that value must necessarily be less certain.
This tradeoff implies some degree of unavoidable uncertainty.
Reproducible, reliable inference will likely require development of new inference methods
that explicitly model continuous geography.
